\documentclass[a4paper]{article}

\title{Solutions to Practice Exam~1}
\author{L. F. \textsc{Jaw}}

\input{preamble-all}
\input{preamble-en}
\input{preamble-begin}

\begin{document}
\maketitle

This is my own solution to Practice Exam~1 of
\href{https://ocw.mit.edu/courses/mathematics/18-014-calculus-with-theory-fall-2010/exams/}{18.014}.

\begin{enumerate}
\item Compute \(\int_{99}^{103} \paren{2x-198}^2 [\sqrt{x-99}] \dx\)
  where here \([x]\) is defined to be the largest integer \(\le x\).

  We have
  \begin{align*}
    \int_{99}^{103} \paren{2x-198}^2 [\sqrt{x-99}] \dx
      &= \int_0^4 \paren{2x}^2 [\sqrt x] \dx \\
      &= \int_0^1 0 \dx + 4 \int_1^4 x^2 \dx \\
      &= 4 \cdot \frac{4^3 - 1^3}{3} \\
      &= 84.
  \end{align*}

\item Let \(S\) be a square pyramid with base area \(r^2\) and height
  \(h\).  Using Cavalieri's Theorem, determine the volume of the
  pyramid.

  By elementary geometry, we have
  \[
    a_S(x) = \frac{(h-x)^2 r^2}{h^2}.
  \]
  Then
  \begin{align*}
    v(S) &= \int_0^h a_S(x) \dx \\
         &= \int_0^h \frac{(h-x)^2 r^2}{h^2} \dx \\
         &= \frac{r^2}{h^2} \int_0^h (h-x)^2 \dx \\
         &= \frac{r^2}{h^2} \int_{-h}^0 x^2 \dx \\
         &= \frac{r^2}{h^2} \cdot \frac{h^3}{3}
           = \frac{r^2 h}{3}.
  \end{align*}

\item Let \(f\) be an integrable function on \([0, 1]\).  Prove that
  \(\abs{\,f}\) is integrable on \([0, 1]\).

  \begin{proof}
    Since \(f\) is integrable on \([0, 1]\), for any \(\epsilon > 0\), we
    can always find step functions \(s\) and \(t\) such that
    \(s \le f \le t\) and
    \[
      \int_0^1 t - \int_0^1 s < \epsilon.
    \]
    Let \(P = \brce{x_0, x_1, \dotsc, x_n}\) be a common partition for \(s\)
    and \(t\).  Denote \(s(x) = s_k\) and \(t(x) = t_k\) on every open
    interval \((x_{k-1}, x_k)\).  We construct two other step functions
    \begin{gather*}
      s'(x) =
      \begin{cases}
        s(x), & s(x) \ge 0, \\
        -t(x), & t(x) \le 0,
      \end{cases}
      \qand
      t'(x) =
      \begin{cases}
        t(x), & s(x) \ge 0, \\
        -s(x), & t(x) \le 0.
      \end{cases}
    \end{gather*}
    It is easy to verify that \(s' \le \abs{\,f} \le t'\) and
    \[
      \int_0^1 t' - \int_0^1 s' = \int_0^1 t - \int_0^1 s < \epsilon.
    \]
    This means that \(\abs{\,f}\) also satisfies the Riemann condition and
    is thus integrable.
  \end{proof}

\item The well-ordering principle states that every non-empty subset of
  the natural numbers has a least element.  Prove the well-ordering
  principle implies the principle of mathematical induction.

  \begin{proof}
    Let \(T\) be the set of all positive integers not in \(S\).  Suppose
    \(T\) is nonempty.  By the well-ordering principle, \(T\) has a smallest
    member.  Let \(m\) denote such a member.  The fact that \(m \notin S \)
    implies \(m \ne 1\).  It cannot be the case that \(m < 1\) since \(1\)
    is the smallest positive integer.  Thus, \(m > 1\).  Then \(m-1\) is
    another integer since \(\mathbf{Z}\) is closed under subtraction.  And
    it is positive since \(m > 1\) implies \(m-1>0\).  It cannot be in
    \(S\).  If so, then \(m\) will be in \(S\), which contradicts our
    definition of \(T\).  This means \(m-1\) is another positive integer not
    in \(S\) and is thus in \(T\).  Notice that \(m-1\) is smaller than
    \(m\).  This contradicts the well-ordering principle.  Thus, \(T\) must
    be empty.  This is to say that \(S\) contains all positive integers.
  \end{proof}

\item Suppose \(\lim_{x \to p^+} f(x) = \lim_{x \to p^-} f(x) = A\).
  Prove \(\lim_{x \to p} f(x) = A\).

  \begin{proof}
    For any \(\epsilon > 0\), there exist \(\delta_1, \delta_2 > 0\) such
    that
    \begin{gather*}
      \abs{\,f(x) - A} < \epsilon \txt{whenener} 0 < x - p < \delta_1 \\
      \iand
      \abs{\,f(x) - A} < \epsilon \txt{whenener} 0 < p - x < \delta_2.
    \end{gather*}
    Let \(\delta = \min\brce{\delta_1, \delta_2}\).  Then
    \[
      \abs{\,f(x) - A} < \epsilon \txt{whenener} 0 < \abs{x - p} < \delta.  \qedhere
    \]
  \end{proof}
\end{enumerate}
\end{document}
