\documentclass[a4paper]{article}

\title{Solutions to Problem Set 1}
\author{L. F. \textsc{Jaw}}

\input{preamble-all}
\input{preamble-en}

\renewcommand{\thefootnote}{\fnsymbol{footnote}}

\input{preamble-begin}

\begin{document}
\maketitle

This is my own solution to the problem set 1 of
\href{https://ocw.mit.edu/courses/mathematics/18-014-calculus-with-theory-fall-2010/assignments/}{18.014}.
The first three problems are from Apostol's \emph{Calculus} and the next
three problems are from Munkres's Course Note A (1: 18, 20, 43; 9--10).

\begin{enumerate}
\item Prove Theorem I.11: If \(ab=0\), then \(a=0\) or \(b=0\).

  \begin{proof}
    If not, then \(a \ne 0\) and \(b \ne 0\), which implies
    \begin{align*}
      1 &= 1 \cdot 1           && \reason{by the identity axiom,} \\
        &= (aa^{-1}) (bb^{-1}) && \reason{by the inverse axiom,} \\
        &= (a^{-1}a) (bb^{-1}) && \reason{by commutativity,} \\
        &= a (a(bb^{-1}))      && \reason{by associativity,} \\
        &= a ((ab)b^{-1})      && \reason{by associativity,} \\
        &= a (0 \cdot b^{-1})  && \reason{by assumption,} \\
        &= a \cdot 0           && \reason{by Theorem I.6,} \\
        &= 0                   && \reason{by Theorem I.6.}
    \end{align*}
    But we know by Axiom 4 that \(1 \ne 0\).  Thus, either \(a = 0\) or
    \(b = 0\).
  \end{proof}

\item Prove Theorem I.25: If \(a < c\) and \(b < d\), then \(a+b < c+d\).

  \begin{proof}
    We have
    \begin{align*}
      a + b &< c + b && \reason{by Theorem I.18,} \\
            &= b + c && \reason{by commutativity,} \\
            &< d + c && \reason{by Theorem I.18,} \\
            &= c + d && \reason{by commutativity.} \qedhere
    \end{align*}
  \end{proof}

\item Prove that \(\abs*{\abs x - \abs y} \le \abs{x - y}\).

  There are two ways to prove this.  One is logical and the other is
  algebraic.

  \begin{proof}
    If \(\abs x - \abs y \ge 0\), then
    \[
      \abs*{\abs x - \abs y} = \abs x - \abs y \le \abs{x - y}.
    \]
    If \(\abs x - \abs y < 0\), then
    \[
      \abs*{\abs x - \abs y} = - (\abs x - \abs y) = \abs y - \abs x \le \abs{y - x} = \abs{x - y}.
    \]
    In all cases, the inequality holds.
  \end{proof}

  \begin{proof}
    We have
    \begin{align*}
      xy &\le \abs{xy} = \abs x \abs y         && \iff \\
      -2\abs x\abs y
         &\le -2xy                             && \iff \\
      \abs x^2 - 2\abs x\abs y + \abs y^2
         &\le x^2 - 2xy + y^2                  && \iff \\
      \paren{\abs x - \abs y}^2
         &\le \paren{x - y}^2                  && \iff \\
      \sqrt{\paren{\abs x - \abs y}^2}
         &\le \sqrt{\paren{x - y}^2}           && \iff \\
      \abs*{\abs x - \abs y} &\le \abs{x - y}. && \qedhere
    \end{align*}
  \end{proof}

\item Prove Theorem 6: If \(a\) and \(b\) are in
  \(\mathbf{P}\), so is \(ab\).

  \begin{proof}
    We choose a fixed positive integer \(a\) and then prove this theorem
    by induction on \(b\).  First, \(a \cdot 1\) is in \(\mathbf{P}\),
    since \(a \cdot 1 = a\) by the identity axiom and \(a\) is in
    \(\mathbf{P}\).  Now suppose \(b\) is a positive integer such that
    \(ab\) is in \(\mathbf{P}\).  Then we have \(a(b+1) = ab + a \) by
    distributivity.  By the induction hypothesis, \(ab\) is a positive
    integer.  And \(a\) is also a positive integer.  We have already shown
    that \(\mathbf{P}\) is closed under addition.  Thus, \(ab + a\) is
    also a positive integer, which means \(a(b+1)\) is in \(\mathbf{P}\).
    This proves that \(\mathbf{P}\) is also closed under multiplication.
  \end{proof}

\item Prove Theorem 12: \(a^n b^n = \paren{ab}^n\), where \(a\) and
  \(b\) are any real numbers and \(n\) is a positive integer.

  \begin{proof}
    The formula is trivially true for \(n=1\), since
    \(a^1 b^1 = ab = \paren{ab}^n\).  Now suppose the formula holds
    for some \(n\).  Then
    \begin{align*}
      a^{n+1} b^{n+1}
      &= (a^n a) (b^n b) && \reason{by definition of exponent,} \\
      &= (a^n b^n) (ab)  && \reason{by associativity and commutativity,} \\
      &= (ab)^n (ab)     && \reason{by the induction hypothesis,} \\
      &= (ab)^{n+1}      && \reason{by definition of exponent.}
    \end{align*}
    This proves the formula holds for any positive integer \(n\).
  \end{proof}

\item Let \(a\) and \(h\) be real numbers; let \(m\) be a positive
  integer.  Show by induction that if \(a\) and \(a+h\) are positive,
  then
  \[
    (a+h)^m \ge a^m + ma^{m-1}h.
  \]

  \begin{proof}
    The inequality is true for \(m=1\), since
    \(\paren{a+h}^1 = a + h = a^1 + 1 \cdot a^{1-1} h\).  Now suppose it
    is true for some \(m \ge 1\), then
    \begin{align*}
      \paren{a+h}^{m+1}
      &= \paren{a+h}^m \paren{a+h}              && \reason{by definition of exponent,} \\
      &\ge \paren{a^m + ma^{m-1}h} \paren{a+h}  && \reason{by the induction hypothesis} \\
      &                                         && \reason{and positivity of \(a+h\),} \\
      &= a^{m+1} + ma^m h + a^m h + ma^{m-1}h^2 && \reason{by distributivity, associativity,} \\
      &                                         && \reason{commutativity, and definition of exponent,} \\
      &= a^{m+1} + (m+1)a^m h + ma^{m-1}h^2     && \reason{by associativity, the identity axiom,} \\
      &                                         && \reason{and distributivity,} \\
      &\ge a^{m+1} + (m+1)a^m h                 && \reason{by positivity of \(m\), \(a\), and \(h^2\),} \\
      &                                         && \reason{the closure axiom, and Theorem I.18.} \qedhere
    \end{align*}
  \end{proof}

\item[7.\footnotemark] \footnotetext{Only to be attempted once other
    problems are completed} Let
  \[
    A_n = \frac{x_1 + x_2 + \dotsb + x_n}{n} \quad \text{and} \quad
    G_n = \paren{x_1 x_2 \dotsm x_n}^{1/n}
  \]
  represent the arithmetic and geometric mean respectively for a set
  of \(n\) positive real numbers.
  \begin{enumerate}
  \item Prove that \(G_n \le A_n\) for \(n = 2\).

    \begin{proof}
      Let \(a\) and \(b\) be any arbitrary real numbers.  Then
      \begin{align*}
        (a - b)^2 &\ge 0   && \reason{by Theorem I.20, }  && \iff \\
        a^2 + b^2 &\ge 2ab && \reason{by simple algebra.}
      \end{align*}
      By Theorem I.35, \(x_1^{1/2}\) and \(x_2^{1/2}\) are meaningful.
      Let \(a = x_1^{1/2}\) and \(b = x_2^{1/2}\).  Then
      \begin{align*}
        \frac{x_1 + x_2}{2} &\ge \paren{x_1 x_2}^{1/2} && \reason{by simple algebra,} && \iff \\
        G_2 &\le A_2                                   && \reason{by definition.} && \qedhere
      \end{align*}
    \end{proof}

  \item Use induction to show \(G_n \le A_n\) for any \(n = 2^k\)
    where \(k\) is a positive integer.

    \begin{proof}
      The previous part proves that the inequality holds for
      \(k = 1\).  Suppose the inequality holds for some \(k \ge 1\).
      By the induction hypothesis, we have
      \begin{align}
        \label{eq:1}
        \frac{1}{2^k} \sum_{i=1}^{2^k} x_i
          &\ge \paren*{\prod_{i=1}^{2^k} x_i}^{1/2^k} \\
        \intertext{and}
        \label{eq:2}
        \frac{1}{2^k} \mspace{9mu} \sum_{\mathclap{i=2^k+1}}^{2^{k+1}} \mspace{4mu} x_i
          &\ge \paren*{\mspace{12mu} \prod_{\mathclap{i=2^k+1}}^{2^{k+1}} \mspace{5mu} x_i}^{\mathrlap{1/2^k}}.
      \end{align}
      By adding~\eqref{eq:1} and~\eqref{eq:2} and halving the sum, we obtain
      \begin{displaymath}
        \frac{1}{2^{k+1}} \sum_{i=1}^{2^{k+1}} x_i
        \ge \frac12 \brce*{\paren*{\prod_{i=1}^{2^k} x_i}^{\mathrlap{1/2^k}} \mspace{12mu} + \mspace{9mu} \paren*{\mspace{12mu} \prod_{\mathclap{i=2^k+1}}^{2^{k+1}} \mspace{5mu} x_i}^{1/2^k}}
        \ge \paren*{\prod_{i=1}^{2^{k+1}} x_i}^{1/2^{k+1}}
      \end{displaymath}
      The last inequality sign holds by applying the induction hypothesis to
      the right-hand sides of both~\eqref{eq:1} and~\eqref{eq:2}.
    \end{proof}
    
  \item Now for any positive integer \(n\), suppose \(n < 2^m\) for some
    integer \(m\).  Using the set
    \[
      \brce*{x_1, x_2, \dotsc , x_n, A_n, A_n, \dotsc, A_n}
    \]
    where the \(A_n\) appears \(2^m − n\) times in the set, show that
    \(G_n \le A_n\).

    \begin{proof}
      Apply to the set the result from the previous problem and we have
      \begin{align*}
        A_n &= \frac{1}{2^m} \brkt*{\paren{2^m - n}A_n + n A_n} \\
            &= \frac{1}{2^m} \brce*{\paren{2^m - n}A_n + \sum_{i=1}^n x_i} \\
            &\ge \brce*{\paren*{\mspace{6mu} \prod_{i=1}^{\mathclap{2^m-n}} A_n}  \paren*{\prod_{i=1}^{n} x_i}}^{1/2^m} \\
            &= \paren*{A_n^{2^m-n} G_n^n}^{1/2^m} \\
            &= \paren*{\frac{A_n}{G_n}}^{\frac{2^m-n}{2^m}} G_n.
      \end{align*}
      Now suppose \(A_n < G_n\), then we have
      \begin{displaymath}
        \frac{A_n}{G_n} < \paren*{\frac{A_n}{G_n}}^{\frac{2^m-n}{2^m}} < 1,
      \end{displaymath}
      which implies
      \begin{displaymath}
        A_n = \frac{A_n}{G_n} \cdot G_n < \paren*{\frac{A_n}{G_n}}^{\frac{2^m-n}{2^m}} G_n \le A_n.
      \end{displaymath}
      This is absurd.  Thus, it must be the case that \(G_n \le A_n\), which
      is consistent with the above result.
    \end{proof}
  \end{enumerate}
\end{enumerate}

\end{document}
